\part{Introduction}

\textbf{Welcome to} \textbf{\textit{Green Cloaks}}

The \textit{Green Cloaks} game takes place within the Segovax Cluster, a significant star cluster far from Terra. Five systems here form the bastion of humanity on this isolated fringe of the galaxy. The cluster is a place ringed in beauty, mystery and legend. It is here the fabric of reality is said to flex with the ebb and flow of time, and it is amongst these very stars that you will decide the future of humankind and its allies.

The human race, united under the banner of the Terran Sovereignty, has spread throughout the galaxy in the last 3000 years, colonising and terraforming planets to make them habitable, an expansion made necessary following a cataclysmic nuclear war and 1000 years of ruin. They share the galaxy with a multitude of other races, and informal alliances have been formed between them. Some of these races

have the knowledge of technology far more advanced than others; some are pacifist in nature; others seem primitive but with a complex and deep sense of loyalty. Together they fight against their mutual enemies

to push back the front lines from their homeworlds, and in some cases reclaim their lost planets from the aggressor.

Terran technology has advanced little in this time, as humans concentrated instead on spreading out to the stars lest humanity come close to extinction again, and the tech of the enemy is often superior. New tools must be designed, new weapons engineered to allow for the chance of victory.

You are enrolled in one of four regiments that make up a special taskforce in the Terran Sovereignty Army known as the Green Cloaks, named for their ballistic weave cloaks, the only armour that can be mass- produced with the limited resources available at this time of war. Your missions are varied and dangerous, pitted against the enemies of the human race and its allies.

Join the regiments of the Green Cloaks as they seek to defend the Sovereignty, defeat their enemies and overcome the challenges of a seemingly hostile universe.

\textit{Green Cloaks} is a NERF-based LARP system founded in 2012 by Enys Coggles with collaborator Adam Wojcik. As a grassroots system it is perfect for individuals new to LARP wanting to try out the hobby, but it has plenty to also grab the interest of more experienced LARPers and roleplayers. The system offers a friendly atmosphere and an enthralling plot for people of all levels of LARP experience.

With a friendly game team always willing and able to help enhance your experience and gameplay, our aim is to ensure that you are encouraged in your LARP experience, and to bring new people to the hobby.

Thank you for your interest, and we at \textit{Green Cloaks} look forward to seeing you at our events.

The \textit{Green Cloaks} rulebook is divided into five parts, and not all of them are applicable to every character

you may wish to play.

The \textbf{\textit{Introduction}} gives you an explanation of what to expect at a LARP game, and features the timeline of events leading up to very recent years. This timeline is useful to give you a context in which to play the game, and a feel for the world in which \textit{Green Cloaks} is set.

\textbf{\textit{Part 1}} is all about playing the game. This includes rules to keep everybody safe and acceptable behaviour, which we hope you will read at least once. However, this is also the place to look for some of the essential rules of the game, including the \textbf{calls} the game uses to indicate certain effects, how \textbf{dying}, \textbf{healing}

and \textbf{stabilisation} works, how your character's \textbf{hits} work, and what an \textbf{encounter} is. Many of the uses of skills in \textit{Green Cloaks} rely on this concept of an encounter, so you must read this section to gain an understanding of it. Please note that although \textit{Chapter 8: Calls} seems to list a lot of calls, they include all Omega ability calls (analogous to spell effects in other systems), which in many other LARP systems and RPG games would be listed separately. They are listed together so that you have a single section to reference rather than two.

\textbf{\textit{Part 2}} guides you through \textbf{character creation}, including lots of information about the different \textbf{races} you can play, the \textbf{planets} your character could come from, and the \textbf{regiments} you can join. This is also the place where you will find information on all the \textbf{skills} you can use, \textbf{classes} you can play, and \textbf{equipment} your character might use. This is pretty essential reading for any player, and no doubt you'll refer to it frequently as you experience more of the \textit{Green Cloaks} system.

\textbf{\textit{Part 3}} is where you can find information about four very specific areas of the game: \textbf{Engineering}, \textbf{Pharmacology}, \textbf{Thaumaturgy} and the \textbf{Omega plane}. If you don't want to play somebody with the Engineering or Pharmacology skills, or don't want to play a character that uses Omega powers, you can ignore this section (though you might find it interesting to know how this works).

Finally, the \textbf{\textit{Appendices}} are useful resources, filled with handy skill times and costs charts for ease of reference, diagrams, sample character builds, and crafting lists. You'll also find a guide to blaster safety provided by \textit{Blastersmiths UK}, a guide to the categories that currently available blasters fall into, and copies of the character sheet and disclaimer you'll need to fill out at events.

LARP (Live Action Roleplay), is an immersive game that can be described as a mixture of collaborative storytelling, creative world-building, outdoor activity, and sport. It's a game for everyone, regardless of background, profession, or gender: the only requirement is a little bit of imagination (which everybody has). In a LARP game you create a character that lives in, and interacts with, a persistent world that is created and populated by the game organizers, referees and crew, in which there are other players doing exactly the same. Your character will have a personality, skills and abilities, and these will help you decide how to react to this world and what happens in it. You can create almost any kind of character, and that character has the chance to become a hero, earn a reputation as a villain, be a mad scientist or hard-working medic, excel as a warrior or investigate the hidden secrets of the world. Whatever character you choose to play is up to you, and the limit is your imagination.

LARP is also a hobby that has very passionate players, who demand engaging events, exciting storylines, and opportunity for character development. However, the amount of time or resources you put into the game is up to you. Players can spend time choosing or creating kit (armour, costumes, accessories) to suit them and their character's personality, skills or profession; they may spend time practising combat with LARP-safe weapons; or they may spend time outside of events writing stories and songs to tell at the next event. LARP events are also often populated by traders who make LARP-safe weapons and armour, and who are just as passionate about the game as the players. Between the players, traders, game organizers, crew and referees, and entire world is collectively created and the stories of that world come to life.

A LARP game can be set in any world or setting (also called a system) and as a character you have the opportunity (or challenge) to fit into that world as a person or creature. When the game begins you are then let loose to explore the world around you and dive into it. From there it's up to you to explore and react to what happens. The Game Team (referees, monsters and other crew) will work hard to bring the

environmental and physical elements of that world to life. How you react to the events in front of and around you will directly affect you and other characters.

There is a combat element to most LARP games, in which LARP-safe weapons (carbon fibre and fibreglass cores encased in foam and painted latex, that look like real weapons but which are soft and safe to use) and/or foam dart guns are used. Small skirmishes, camp attacks, or large-scale field battles are played out during most games, and for many players these battles are the highlight of an event.

LARP can be a confidence-building activity for those who participate, and it is also a great way to meet people and make lifelong friends: after all, when you've faced down with somebody the challenges and terrors that the game organizers create, you have a strong foundation for a friendship!

We hope you that enjoy LARP as much as we, and the thousands of others already playing, do. Good luck, and have fun!

\textbf{The} \textbf{\textit{Green Cloaks}} \textbf{Timeline}

\textbf{2209}: Nuclear war on Earth with cataclysmic results.

\textbf{Circa 3300}: After over a thousand years of conflict and ruin, global civilisation is re-established as a

\textit{de facto} sovereignty under its first king, Michael I. Emphasis is placed on rebuilding and searching for resources beyond those of Earth. Earth is renamed Terra in reference to great, ancient civilisations, and as a sign of new beginnings.

\textbf{3500-3521}: Humanity establishes the first colonies on the moon, Mars and Titan. Rich resource deposits are exploited and the resources shipped to Terra. Much of the previous infrastructure and knowledge has been lost, however, and some doubt if humanity has recovered enough of what it once knew. With the advent of space travel more effort is put into expansion rather than research.

\textbf{3552}: Sleeper ships are sent out to investigate potential habitable planets. Thousands of hopeful colonists are frozen and flown in sublight engines into the void. Most are lost or destroyed, however a few make it to their destinations and form the first human colonies outside Sol.

\textbf{3750}: Humanity's first contact with an alien race. Sleeper ship 2339 reaches its destination to find a thriving pre-industrial civilisation. After an accidentally hostile first encounter the soldiers guarding the expedition decide that use of force is required, and nuclear weapons are deployed.

\textbf{3811}: Physicists make a breakthrough in the manipulation of space-time and develop the singularity drive, or ``S-drive''. This allows travel over great distances at a much faster speed than before.

\textbf{3850}: Many nearby star systems are explored and planets are marked out as potential targets for terraforming, the process which should make viable planets habitable for human life.

\textbf{3873}: Sleeper ship 2339 is discovered; the planet it landed on is a radioactive husk. No life remains. Logs are located in the ship detailing the attack on the natives and subsequent fusion of unknown radioactive particles, resulting in the ignition of the atmosphere.

\textbf{3875}: Despite an attempted cover-up, the bulk of humanity learns of the first contact with an intelligent alien race and its subsequent destruction. The next 10 years sees mass civil unrest over colonisation policies and the role of the military in them.

\textbf{3900-3990}: Mars is terraformed, allowing citizens to live outside of the habitation domes they were

previously confined to. Several planets undergoing terraforming fail and are abandoned.

\textbf{4000}: First contact made with the vrede. Initial distrust is quickly swept aside as they appear pacifist and willing to share their highly advanced technology. Humanity's technology advances many hundreds of years in a very short space of time; terraforming technology becomes relatively easy, and many planets are colonised as a result.

\textbf{4006}: Knowledge of the existence of the vrede is made public. Most of humanity regards this as the beginning of a new age of peace and opportunity. For some, however, learning that humanity is not the only star-faring race causes worry and doubt.

\textbf{4119}: Civil unrest breaks out. A faction known as Humanity First becomes popular, and many worlds are thrown into political turmoil. They preach that the vrede are not to be trusted, stating that humanity still knows almost nothing about them and that no race as advanced as they are would be without weapons.

\textbf{4150}: Humanity First gains traction on frontier worlds near the vrede border. Three sections of the Terran military stationed in the area side with them and declare they will assault vrede space. Forces are dispatched from Terra to prevent this from leading to full-scale civil war. Other planets join Humanity First's cause after the Terran forces attack. No attack is ever launched against the vrede.

\textbf{4600}: After 450 years of civil war and billions of lost lives, the last planet held by Humanity First falls.

The organisation is officially banned. The Terran Sovereignty has been set back hundreds of years by the internal conflict.

\textbf{4797}: Colonisation of the Segovax cluster, approximately 80 years travel from Terra, begins. It has an unusually high amount of viable planets for terraforming, as well as an abundance of natural resources. Some planets already appear habitable after some initial samples are taken. A planet that would later become Delmont is colonised and used as a staging area, becoming a hub for settlers and trading. Its population skyrockets due to immigration.

\textbf{4800}: Report from scouts sent deeper into the cluster tell of strange creatures and monsters as well as other highly strange occurrences. These reports, however, are largely dismissed by colonisation command, and repressed fearing civil unrest.

\textbf{4805}: Cantiacorum, which would later become the capital of the Segovax cluster, is colonised. It is perfect for human life and becomes the political hub for the area.

\textbf{4859}: Marazion V, with its abundance of valuable resources, is colonised. Subterranean cities are built to allow for further expansion on this mountainous planet. A few surface citadels are built to act as spaceports and trade hubs.

\textbf{4865}: Ardheim, with its hostile forests and dangerous wild beasts, is finally colonised. Given the previous destruction of natural habitats, great care is taken by the colonists to preserve the heritage of the planet. Great walled cities are built into the forest including its capital city of Kingskeep.

\textbf{4868}: Tetrarch, a desert planet that would later be the sole manufacturer and exporter of silicon, is discovered and initially used as an interplanetary scrapyard by neighbouring colonies.

\textbf{4889}: Durgan, a planet with small continents and archipelagos is discovered. The presence of an atmosphere similar to that of Terra, with warm oceans, requires very few alterations, and it is colonised with relative ease. Hindered only by the lack of sea-going vessels, it is fully established within the next decade.

\textbf{4899}: Rossi, a small planet located in the same solar system as Ardheim, is colonised. Rich in mineral wealth but with a harsh, freezing environment, some terraforming is required. The Rossii people, with the assistance of the people of Ardheim, are quick to adapt, and the planet is soon an industrial and shipping hub of the sector.

\textbf{4910}: Pushing further into the Segovax cluster a primitive yet intelligent species, who call themselves the mascen, are discovered. They are quick to adopt friendly relations with the Terran Sovereignty. The mascen hold honour in high regard and they return the gesture of respect and friendship given to them by humans.

\textbf{4950}: Colonies on distant worlds are suddenly attacked by unknown hostile forces of lizard-like creatures later identified as tae'go. Their raids are mainly focused on stealing food and supplies, though the need for these raids is not discovered for some time.

\textbf{4951}: Contact is lost with a distant, recently colonised cluster of worlds, F4R-ZQ; breakdown in communication is thought to be due to ion storms in a nearby nebula. Scouts are dispatched to re- establish contact.

\textbf{4996}: The scout ships are obliterated by an unknown hostile fleet.

\textbf{4997}: The unknown fleet reaches the borders of Sovereignty space and issues a broadcast across the

cluster: ``We are the One Bakkar, surrender and join us. Resist and you shall be destroyed''.

\textbf{4998}: The first ground skirmish takes place with the One Bakkar on the planet Vero. Information from the planet records the defenders seeing a highly varied mix of alien races, some of which they had never encountered before. Oddly, however, the main bulk of the force seem to be human. The One

Bakkar fight using fanatical shock troops that throw themselves against the human front line, whittling them down with sheer weight of numbers. Before long the defenders are overwhelmed and Vero is lost. The planet's fate remains unknown to this day but it is most likely a barren wasteland, stripped of resources to fuel the One Bakkar crusade.

\textbf{4999}: All-out war breaks out along the edges of Sovereignty space as the One Bakkar attack border colonies. Many colonies are lost and victories are few and short-lived. The One Bakkar have the advantage in both numbers and firepower. Some colonies simply surrender, their citizens then being deported as troops in the main One Bakkar force in later engagements. A few of the Terran conscripts are recaptured and interrogated: all refuse to recant and seem to be totally loyal to their new masters. A form of advanced indoctrination is suspected. A petition to the vrede for aid is made but refused.

\textbf{5001}: In the chaos of the outbreak of war many mascen are enslaved by corrupt Terran corporations for their ingenuity and strength. Even if morally repugnant, this action arguably saves the species, as the mascen homeworld is soon targeted and seized by the One Bakkar. It remains under enemy control to this day, located on the far side of the cluster well beyond Sovereignty space.

\textbf{5001-5009}: As the fight continues against the One Bakker the Sovereignty learns more about their attackers. They are a collective of alien species with a cult-like mentality. It is unknown which of these species is the founder, and what their overall motives are. It is known, however, that they expand their numbers with a highly effective and rapid method of indoctrination that is not fully understood. After several years of fighting it seems that the vast majority of the One Bakkar troops are former citizens.

\textbf{5010}: The One Bakkar continue to advance, their numbers growing with time. It continues to push back the Terran Sovereignty on all fronts.

\textbf{5011}: The One Bakkar fleets employ new blitzkrieg tactics and make a beeline for Terra. The key trade routes to Terra which have acted as a lifeline to less-established planets are shattered. Many lose contact with the Terran Sovereignty completely; some are cut off and destroyed by the One Bakkar, while some are ignored as the One Bakkar make their push for Terra. The main supply planets of the Segovax cluster are cut off and form a coalition to defend their home systems. Most of the outlying colonies in this area are destroyed. Some are not attacked, however their inhabitants are forced to fend for themselves for many years. Since most colonised planets are reliant upon Terran trade links, many colonies die out. Some surrender to the invaders and quickly join their ranks to continue the momentum of the attack.

\textbf{5012}: The core planets in the Segovax cluster band together to resist the One Bakkar push and the crippling loss of supplies from Terra. Ardheim, Cantiacorum, Delmont, Durgan, Marazion V, Rossi and Tetrarch keep communication, trade and resistance alive in the area. No planet can fully survive on its own and resources are transported between systems to keep all the planets alive. The first combined military of these planets is formed, and works together to keep the trade routes open and the space around these planets secure. This force is known as the ``Green Cloaks'', named for the ballistic weave cloaks worn by its soldiers, the only armour that can be mass-produced with the limited resources available at this time of war. The One Bakkar push them hard, however their main force carries on towards Terra, leaving a significant force behind to keep the crippled planets fenced in.

\textbf{5256}: After hundreds of years the relentless One Bakkar have pushed almost as far as Terra. The Terran Sovereignty begins to give up hope. The main thrust of the One Bakkar attacks include their crack troops, and they are victorious time and time again.

\textbf{5324}: Just as the One Bakkar are about to descend upon Terra a massive vrede fleet comes out of nowhere and obliterates the main One Bakkar assault in a matter of hours. The fact the vrede have been able to deploy such power and kept it quiet all this time causes much distrust toward them among Terrans. The fleet is called back to vrede space almost immediately and the Terrans are left to reinforce their lines as best they can. When questioned about this fleet the vrede only ever refuse to comment and many younger vrede seem to be surprised at hearing of its existence.

\textbf{5325-5600}: The Terran Sovereignty breaks the blockade of many border colonies including that of the

Segovax cluster which is finally reached in 5577. It becomes the new front line in Terran space. Civil wars break out among colonies that felt the Terran Sovereignty had abandoned them. Several try to declare independence. Some are even tenacious enough to succeed. These places become havens for criminals and outlaws running from Terran Sovereignty law. The One Bakkar are rarely seen and many suspect they have been defeated, however reports from the front line indicate that they are gathering their strength after their main fleet was destroyed. The mascen are freed from slavery and their

captors face severe penalties, including execution in extreme cases. The mascen are offered a place among Terran society which they gladly accept. Many join the military with a view to taking back their homeworld.

\textbf{5700}: The tae'go approach the Terran Sovereignty on peaceful terms claiming they have information about the One Bakkar. They had been living on Terran-occupied planets for centuries while hiding from them after their homeworld, Tazrak, was captured and they were forced to flee. They offer their services if the Terrans give their old, sick and young a safe place to live. Having faced their unpredictable attacks before, the Terrans agree and many tae'go enter service as scouts, guides and in some cases assassins.

\textbf{5800}: After many years of tense standoff and only the occasional skirmish with One Bakkar scouts, humanity starts to feel safe again. They are contacted by the enigmatic myr'na who seek to make peace with the Terran Sovereignty. The myr'na are at a turning point in their history, with many young myr'na wanting to explore the galaxy and many older myr'na wishing to stick to their old ways. The myr'na are secretive about their power.

\textbf{6012}1: After many years of strained peace it seems the One Bakkar are ready to renew their assault on the Terrans. The planet Zennor in the Segovax cluster comes under heavy attack. The Green Cloaks are deployed to repel the ground assault. Patient 0 of the Adept Program is found, sparking a new understanding of the Omega plane, previously thought a myth (outside of the colonies located deep within the Segovax cluster). The Adept Program begins.

601$2 = 2012. $The \textit{Green Cloaks} system began in 2012. For timeline events that occurred after 6012, please refer to the website.
